Analysis

### 2024-12-26 production/prod_analysis-fix-t5.Rmd
changes methods + analisis

# Df variables nivel 1 _________________________________________________________
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(ola,
                                               just_educ, des_perc, merit_talent, merit_effort,
                                               educ, quintil1:sexo)

@@

for (i in c("just_educ", "des_perc",
            "merit_effort","merit_talent")) {
  df_study1[[i]] <- factor(df_study1[[i]],
                            levels = 1:5,
                            labels = c("Strongly disagree","Disagree",
                                       "Neither disagree nor agree",
                                       "Agree", "Strongly agree"))
}

@@

#
df_study1$des_perc <- 
sjlabelled::set_label(x = df_study1$des_perc, 
                      label = "Income differences are too large")

@@

df<- dfSummary(x = dplyr::select(df_study1, ola, merit_talent, merit_effort, des_perc, educ, quintil1, ess, edad, sexo, pos_id),
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = TRUE,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = T, # plot
               valid.col = F, # n valido
               col.widths = c(700,10,10,10)
               )
df$Variable <- NULL # delete variable column
summarytools::view(df,
     file = here::here("output/tables/desc02.html")
     ) # Ver tabla en un archivo HTML
webshot::webshot(url =here::here("output/tables/desc02.html"),
        file =here::here("output/tables/desc02.png")) # Remover archivo HTML
# print(df, method = "render")
```

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

```{r alluvial}
colors<- c("#f1eef6ff","#bdc9e1ff","#b3b3b3ff","#74a9cfff","#0570b0ff")
alluvial_justeduc <- ggplot(datos.justeduc, aes(x = ola, fill = just_educ, stratum = just_educ,
                             alluvium = idencuesta, y = porcentaje))+
    ggalluvial::geom_flow(alpha = .66) + 
    ggalluvial::geom_stratum(linetype = 0) +
    scale_y_continuous(labels = scales::percent) + 
    ylab(label = NULL) +
    xlab(label = NULL) + 
    theme(#legend.position = 'top',
          legend.title = element_blank(),
          plot.caption = element_text(hjust = 1)) +
    scale_fill_manual(values = colors) +
    geom_shadowtext(data = etiquetas.justeduc, 
              aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
              position = position_stack(vjust = .5),
              show.legend = FALSE,
              size = 4,
              color = rep('white'),
              bg.colour='grey30')+
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022, 2023)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 9519)")

@@

```{r bivariate-plots}
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
merit_plot <- df_study1 %>% 
  select(merit_effort, merit_talent, des_perc, just_educ) %>% 
  sjPlot::plot_stackfrq(weight.by = df_study1$ponderador_long_total,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=rep(c("bottom", "top", "bottom", "top", "bottom"),4),
                      legend.labels = c("Strongly disagree", "Disagree", "Neither agree
nor disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Merit effort", "Merit talent", "Inequality 
perception", "Justification 
of 
educational inequality")
                      ) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 9519)")
merit_plot

@@

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

# Añadir etiquetas personalizadas
years_plot <- years_plot +
  geom_point(data = label_data, aes(label_x + 1, label_y, color = variable, shape = variable), size = 4) +  # Capa de puntos para los símbolos de forma
  geom_text(data = label_data, aes(label_x, label_y, label = variable), color = "black", hjust = 0, nudge_x = 1.1, nudge_y = 0, size = 4)  +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 9519)")

@@

```{r cor-plot, fig.dim=c(10,6)}
rm(list=ls())
if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               sjPlot,
               sjmisc,
               summarytools,
               lavaan
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
# Full sample_______________________________________________________________________
mat_full<- 
df_study1 %>%
  # filter(ola==1) %>%
  select(just_educ, # distributive justice
         des_perc,
         merit_effort, merit_talent # meritocracy
         ) %>% 
  psych::mixedCor()

@@

from ELSOC 2016-2023 (n = 9519)", x = 0.01, y = 0.05, just = "left", gp = grid::gpar(fontsize = 10, col = "black"))

>>>>

# Df variables nivel 1 _________________________________________________________
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(ola, idencuesta,
                                               just_educ, covid_economia, covid_beneficios, des_perc, merit_talent, merit_effort,
                                               educ, quintil1:sexo)

ids <- 
  df_study1 %>% 
  select(idencuesta,ola) %>% 
  filter(ola==1) %>% 
  sjmisc::frq(idencuesta,show.na = F) %>% as.data.frame()

# filter data by the idencuesta of t7
df_study1 <- 
  df_study1 %>%
  filter(idencuesta %in% ids$val)

@@

for (i in c("just_educ", "des_perc", "covid_economia",
            "merit_effort","merit_talent")) {
  df_study1[[i]] <- factor(df_study1[[i]],
                            levels = 1:5,
                            labels = c("Strongly disagree","Disagree",
                                       "Neither disagree nor agree",
                                       "Agree", "Strongly agree"))
}

df_study1$covid_beneficios<- car::recode(df_study1$covid_beneficios, c("1=2; 2=1"))
df_study1$covid_beneficios <- factor(df_study1$covid_beneficios,
                                     levels = 1:2,
                                     labels = c("No", "Yes"))

@@

df_study1$covid_economia <- 
sjlabelled::set_label(x = df_study1$covid_economia, 
                      label = "It is more important to give priority to economic activity than to the health of the population")

df_study1$covid_beneficios <- 
sjlabelled::set_label(x = df_study1$covid_beneficios, 
                      label = "Obtained state benefits during the covid-19 pandemic")
#
df_study1$des_perc <- 
sjlabelled::set_label(x = df_study1$des_perc, 
                      label = "Income differences are too large")

@@

df<- dfSummary(x = dplyr::select(df_study1, ola, covid_economia, covid_beneficios, merit_talent, merit_effort, des_perc, educ, quintil1, ess, edad, sexo, pos_id),
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = TRUE,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = T, # plot
               valid.col = F, # n valido
               col.widths = c(700,10,10,10)
               )
df$Variable <- NULL # delete variable column
summarytools::view(df,
     file = here::here("output/tables/desc02.html")
     ) # Ver tabla en un archivo HTML
webshot::webshot(url =here::here("output/tables/desc02.html"),
        file =here::here("output/tables/desc02.png")) # Remover archivo HTML
# print(df, method = "render")
```

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,covid_economia, covid_beneficios,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

```{r alluvial}
colors<- c("#f1eef6ff","#bdc9e1ff","#b3b3b3ff","#74a9cfff","#0570b0ff")
alluvial_justeduc <- ggplot(datos.justeduc, aes(x = ola, fill = just_educ, stratum = just_educ,
                             alluvium = idencuesta, y = porcentaje))+
    ggalluvial::geom_flow(alpha = .66) + 
    ggalluvial::geom_stratum(linetype = 0) +
    scale_y_continuous(labels = scales::percent) + 
    ylab(label = NULL) +
    xlab(label = NULL) + 
    theme(#legend.position = 'top',
          legend.title = element_blank(),
          plot.caption = element_text(hjust = 1)) +
    scale_fill_manual(values = colors) +
    geom_shadowtext(data = etiquetas.justeduc, 
              aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
              position = position_stack(vjust = .5),
              show.legend = FALSE,
              size = 4,
              color = rep('white'),
              bg.colour='grey30')+
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022, 2023)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 7894)")

@@

```{r bivariate-plots}
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent, covid_economia, covid_beneficios,
                                               educ, quintil1:sexo) %>% na.omit()
merit_plot <- df_study1 %>% 
  select(merit_effort, merit_talent, des_perc, just_educ) %>% 
  sjPlot::plot_stackfrq(weight.by = df_study1$ponderador_long_total,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=rep(c("bottom", "top", "bottom", "top", "bottom"),4),
                      legend.labels = c("Strongly disagree", "Disagree", "Neither agree
nor disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Merit effort", "Merit talent", "Inequality 
perception", "Justification 
of 
educational inequality")
                      ) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 7894)")
merit_plot

@@

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent, covid_economia, covid_beneficios,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

# Añadir etiquetas personalizadas
years_plot <- years_plot +
  geom_point(data = label_data, aes(label_x + 1, label_y, color = variable, shape = variable), size = 4) +  # Capa de puntos para los símbolos de forma
  geom_text(data = label_data, aes(label_x, label_y, label = variable), color = "black", hjust = 0, nudge_x = 1.1, nudge_y = 0, size = 4)  +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 7894)")

@@

```{r cor-plot, fig.dim=c(10,6)}
rm(list=ls())
if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               sjPlot,
               sjmisc,
               summarytools,
               lavaan
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent, covid_economia, covid_beneficios,
                                               educ, quintil1:sexo) %>% na.omit()
# Full sample_______________________________________________________________________
mat_full<- 
df_study1 %>%
  # filter(ola==1) %>%
  select(just_educ, # distributive justice
         des_perc,
         merit_effort, merit_talent # meritocracy
         ) %>% 
  psych::mixedCor()

@@

from ELSOC 2016-2023 (n = 7894)", x = 0.01, y = 0.05, just = "left", gp = grid::gpar(fontsize = 10, col = "black"))
